### 2023-11-26 analysis/paper/paper.Rmd
revising based on comments

---
title: "Preliminary insights into the use of pottery and culinary pratices in southwestern China"
author:
  - Li-Ying Wang:
      email: liyingw@asihp.net
      institute: [AcS]
      correspondence: true
  - Kuei-chen Lin:
      email: kueichenlin@asihp.net
      institute: [AcS]
      correspondence: false
  - Zhiqing Zhou:
      email: zhouzhiQ2000@163.com
      institute: [InsC]
      correspondence: false
institute:
  - AcS: Institute of History and Philology, Academia Sinica 
  - InsC: Chengdu Institute of Cultural Relics and Archaeology
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
    The spread of new crops or terrestrial animals across areas in southwestern China during the Neolithic period is one of the important archaeological issues to understand the process of adoption of new culinary practices in this region. It is widely accepted that rice was introduced into southwest China in 4700 BP at the latest and millets came even earlier. However, to what extent new crops were used by local peoples at that time remains unclear. The excavation of Guijiabao in the Yanyuan basin in southwestern China uncovered abundant materials with well-preserved stratigraphic records and has provided a unique opportunity to understand the local adoption of new crops. To further understand pottery use and food practices at Guijiabao, we investigated organic residues preserved in potsherds using lipid biomarker identification and compounds-specific stable isotope analysis. The results show signals of terrestrial animals in most potsherds and a C~3~ plant source in one potshard, during the Neolithic period. Our case study indicates that hunting might have been the primary means for obtaining subsistence, supplemented by rice/millet farming. This study also demonstrates an approach using organic residue analysis to explore the extent of the adoption of farming during the Neolithic period in southwestern China.

keywords: |
  Southwestern China; Neolithic period; Lipid residue analysis; Stable isotope analysis; Ceramics; Culinary practices
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300)

image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
```

# Introduction
<!-- intro to the topic and research aim-->
Movements of people between geographically distinct areas are usually accompanied by the exchange of goods, technology, or ideas that could have cultural or economic impacts on local populations involved in the interactions [@Burmeister2000]. In prehistoric southwest China, the Hengduan Mountain ranges have served as passages for people to move between the Qinghai-Tibetan Plateau in the north, the Yunnan–Guizhou Plateau in the south, and the Sichuan Basin in the east. The movements in the Hengduan Mountain ranges facilitated not only cultural exchange, but also the spread of new subsistence practices, such as farming and animal domestication [@Tong1986; @Hein2014; @Chen2015]. With more archaeological sites have been found recently along the Hengduan Mountain valleys, researchers are able to dive deeper into local cultural developments and diversity, the timing and routes of exchange network, and the spread of farming practices during Neolithic period across the Hengduan Mountain ranges and neighboring areas [@Hein2014; @Hein2016; @Ma2021; @Lin2022; @Hao2022; @Ma2021].

Despite a long-standing debate on the routes of rice/millet dispersal from northwest China and the middle Yangtze valley to southwestern China [@Guedes2013; @DAlpoim2018; @Liu2022], it is widely accepted that mixed farming was practiced in southwestern China as early as 4700 BP, where the adoption of farming in southwestern Sichuan is around 4700 BP and in western Yunnan is around 4600 BP [@Jiang2016; @YunnanMuseum1981; @Hao2022; @Zhang2010; @Dal2022]. Recent excavations at Guijiabao in the middle Hengduan Mountain ranges bridges the gap between southwestern Sichuan and Yunnan through the finding of millet and rice around 4700 BP [@Huan2022; @Hao2022]. However, it is still not fully understood the extent of the adoption of farming practices by the ancient people in the Hengduan Mountain ranges. Did crop cultivation become the essential part of local subsistence when they were introduced? What other resources were obtained in addition to those grains?

In this study, we focus on archaeological sites

aim to to explore the culinary practices through food residues within ceramic vessels from Guijiabao and Gaoshan sites

Also, the discussion of local subsistence mostly relies on paleobotanical evidence and zooarchaeological assemblages,  The preservation of animal or plant remains could be heavily impacted by environmental conditions.

to explore the culinary practices in the Yanyuan Basin through food residues within ceramic vessels from Guijiabao and Gaoshan sites. We analyzed small samples to test whether those pottery were used for cooking, and if we could detect new crops or species such as rice, millet, or pigs, as food sources when they spread into the Yanyuan Basin. Guijiabao is a long-term occupation settlement with complete temporal sequences from the Neolithic period to the historical period in Yanyuan that have been excavated so far [@Guijiabao2017a; @Guijiabao2020a]. The archaeobotanical evidence at Guijiabao reveals that foxtail millet, broomcorn millet, and rice were present in the Neolithic period around 5000 BP, earlier than previously thought [@Guijiabao2016; @Huan2022], and millet could have already been part of the local diet [@Lin2022]. Although some of these analyses suggest the presence of mixed millet/rice farming [@Huan2022], we still lack direct evidence of culinary practices during this period. To understand the extent of consuming new crops or terrestrial animals at Guijiabao, we ask the questions: What were the culinary practices that can be deduced by the pottery? Can we identify the new crops or terrestrial animals that were incorporated into the local food diets during the Neolithic period? We answer these questions by examining potsherds using organic residue analysis, which allows us to directly explore food cooked or processed in ceramic vessels.

The recent growing application of organic residue analysis demonstrates its usefulness in understanding ancient pottery use [@Kimpe2004; @Mayyas2022], human subsistence strategies, or culinary practices across a region over time [@Kwak2020; @Craig2011; @Suryanarayan2021; @Zhang2022; @Junno2020]. Examining organic residues, particularly lipids, enables archaeologists to investigate ancient food practices through lipid distributions and specific lipid compounds as biomarkers for tracing food origins; for example, millet can be identified through the biomarker miliacin [@Copley2005; @Whelton2021; @Heron2016]. This is possible for prehistoric ceramics due to the protection of clay matrices from microbiological degradation and water leaching, which allows lipid molecules to accumulate and be well preserved in the porous fabric of pottery for over a thousand years [@Evershed1993; @Evershed2008; @Roffet2017]. In this study, we detected and identified lipid residues absorbed into the ceramic matrix that can reveal the original contents cooked or stored in a vessel during its lifetime of use [@Dudd1999; @Miller2020]. Through our case study, we attempt to investigate culinary practices during a period of frequent cultural interactions that accompanied agricultural expansion.

# Geographical and Cultural Background of Guijiabao in Yanyuan Basin

```{r SW-china-map, fig.cap = "The archaeological sites Guijiabao and Gaoshan and other locations in southwestern China that are mentioned in this study. Map data is from naturalearthdata.com"}
library(here)
knitr::include_graphics(here("analysis", "figures", "Sichuan-sites-map.png"))
```

<!--Yanyuan area and Guijiabao site-->
We focus on the Guijiabao site located in the Yanyuan Basin, which is a pull-apart basin in the middle eastern part of the Hengduan Mountain ranges (Figure \@ref(fig:SW-china-map)). The Guijiabao site allows us to understand the early adoption of agriculture because of its uniqueness and rarity: it contains complete cultural components from the Neolithic period to the historical period in that region [@Guijiabao2017a; @Guijiabao2017b]. A survey in 2015 and subsequent excavations from 2016 to 2019 revealed abundant archaeological materials and features, such as semi-subterranean houses, ground houses, post holes, ash pits, ditches, and burials, suggesting that Guijiabao was a settlement site [@Hao2022]. The chronology of the Guijiabao site can be divided into three phases, the Neolithic period (5000-3700 BP), the Bronze Age (3200-2700 BP), and the late North Song to early Ming dynasties (1200-500 BP) based on 19 radiocarbon dates from plant remains and a human bone. For the Neolithic period, it can be further divided into two subphases, 5000-4500 BP and 4500-3700 BP, according to the content of artifact assemblages and ceramic forms and styles that were identified [@Hao2022].

<!--the major finding at GJB-->
The common vessel forms unearthed at the Guijiabao site include jars, urns, bowls, and basins. The representative ceramic styles in the Neolithic period are grayish high-tempered fine ware and reddish brown coarse ware with decorations such as cord-marked and basket-impressed patterns [@Hao2022; @Guijiabao2017a]. Some ceramic decorations present northern styles, and some ceramics with incised geometric patterns and impressed dot patterns resemble those found in western Yunnan, indicative of relationships between the Guijiabao site and sites in Yunnan and the southern part of the Hengduan Mountains [@Guijiabao2017a]. Plant remains reveal that cultivated crops were dominated by foxtail millet (*Setaria italica*) and broomcorn millet (*Panicum miliaceum*) based on flotation samples collected from ash pits and stratigraphic layers that can be traced back to 5000 BP [@Guijiabao2016; @Hao2022]. This corresponds to a C~4~-based diet based on isotopic analysis of four human bones [@Lin2022]. In addition to millet, a smaller number of rice grains (*Oryza sativa*) has also been found in ash pits, where the earliest one dates to 4700 BP [@Hao2022]. A detailed phytolith analysis for soil samples dating to the Neolithic period (5000-3700 BP) further suggests the presence of rice phytoliths throughout the temporal sequences [@Huan2022].

<!--info about Gaoshan-->
To compare the culinary practices in different ecological zones, we examined a sample from the Gaoshan site in the southwestern Chengdu Plain, an alluvial plain in the western Sichuan Basin at the foot of the Qinghai-Tibetan Plateau (Figure \@ref(fig:SW-china-map)). Excavations at the Gaoshan site reveal abundant materials, including ceramics, stone artifacts, animal and plant remains, and archaeological features such as house foundations, walls, burials, ash pits, ditches, and a well [@Gaoshan2017; @He2020; @JiangGaoshan2017; @ZhouGaoshan2015]. According to similar pottery forms and styles, the Gaoshan site can be dated to the early Baodun period, the earliest Neolithic culture (ca. 4700-3700 BP) in the Chengdu Plain [@Gaoshan2017]. The primary subsistence strategy at the Gaoshan site could have been rice cultivation with pig husbandry, supplemented by millet cultivation [@JiangGaoshan2017; @Lee2020; @He2020]. It is suggested that the spread of rice farming into the Sichuan area occurred at least 4700 years ago from people moving from the Middle and Lower Yangzi valleys to the Sichuan area [@Zhang2010].

Although the recent findings at the Guijiabao and Gaoshan sites support mixed agriculture as the primary subsistence strategy, it is still unclear how much millet or rice was included in the local diets during the Neolithic period. The macro-botanical and phytolith analyses of the Guijiabao site show that millet was the major crop, similar to the findings at the Yingpanshan site in Sichuan [@Zhao2011]. However, plant remains from many sites in southwestern China reveal a predominance of rice, such as at the Gaoshan and Baodun sites in Sichuan [@JiangGaoshan2017] and Baiyangcun sites in Yunnan [@Jiang2016; @Henglanshan2014]. This indicates that there were variations in agricultural practices in southwestern China. To better understand the extent of the adoption of new crops, organic residue analysis of ceramics is a useful approach since we can directly study food residues related to culinary practices, including food preparation, processing, and cooking. This provides another way to examine the expansion of agriculture in the Yanyuan area.

# Materials and methods

## Samples
As a preliminary research to accesses the feasibility of this approach, we selected ceramic fragments that were identified as typical from the Guijiabao and Gaoshan sites. The fragments date to the Neolithic (around 4500 BP) and historical (around 1000 BP) (Table \@ref(tab:sherds-table)), which were determined from known contexts based on radiocarbon dating. Ceramic fragments from the Guijiabao site were excavated from trench 13 excavated during the 2017 field season, along with one fragment from the 2016 season. The pottery samples are mostly gray base fragments with fine sand temper, of which a few have impressed decorations. Although their original forms could not be recovered, the evidence of chars on some fragments and their contexts in cultural layers suggest that these ceramics were used in daily life [@Guijiabao2017a; @Hao2022]. This research also attempts to confirm that these pottery vessels were used as cooking/storage wares for food.

```{r sherds-table, cache= TRUE, cache.extra=tools::md5sum(here::here("analysis/data/raw_data/SW-china-potsherd.xlsx"))}
Ch_sherds <-
  readxl::read_excel(here::here("analysis", "data", "raw_data", "SW-china-potsherd.xlsx"), col_names = TRUE)

knitr::kable(Ch_sherds,
             caption = "A list of the pottery sherds that were analyzed in this study. The sherd labeled 'SYG' indicates potsherds from the Guijiabao site, while the label 'CDG' refers to those from the Gaoshan site")
```

## Methods
We detected and identified lipid compounds using gas chromatography mass spectrometry (GC-MS) and quantified lipids using gas chromatography with flame-ionization detection (GC-FID). To distinguish the source of the common fatty acids, palmitic (C~16:0~) and stearic (C~18:0~), which are found in most animals and plants, we conducted compound-specific isotope analysis to identify stable carbon isotope characteristics using gas chromatography combustion isotope ratio mass spectrometry (GC-C-IRMS). The stable carbon isotope ratio for individual lipid compounds is a powerful indicator of the origins of fatty acids since different food categories, such as terrestrial animals, marine animals, and plants, have specific ranges of δ^13^C values [@Craig2007; @Evershed1994].

For sample processing, the surface of each potsherd was removed to prevent external contamination from environmental factors and handling during excavation or post-excavation analysis. Approximately 3.5 grams of powdered samples were extracted with a mixture of dichloromethane/methanol solution (9:1 v/v) using accelerated solvent extraction (ASE Dionex 350; at 100 °C) [@Mcclure2018]. To recover free fatty acids from bound lipids for a higher yield, we conducted saponification by adding 10 mL of 1M KOH in methanol and 1 mL water to the extracts and heating them in an oven at 70 °C overnight [@Kaluzna2016; @Craig2004]. Saponified extracts were acidified with 6M hydrochloric acid (HCl), and the acid fraction was extracted using hexane ($3\times10$ mL). A 5% aliquot of each sample was derivatized using 20 uL N,O-bis(trimethylsilyl)trifluoroacetamide (BSTFA) with 1% v/v trimethylchlorosilane (TMCS), and heated at 70 °C for one hour. After drying under a stream of nitrogen, the derivatized sample was dissolved in toluene prior to GC-MS.

The remaining 95% aliquot was fractioned using solid phase extraction (Aminopropyl, NH2, column with 0.5 g sorbent) with 8 mL dichloromethane:Isopropyl alcohol (3:1), 8 mL 4% acetic acid in ethyl ether, and 6 mL methanol. The second fraction was methylated to produce fatty acid methyl esters (FAMEs) with 1 mL dry hexane and 2 mL 10% (v/v) acetyl chloride in dry methanol, and heated at 70 °C overnight. The FAME derivatives were extracted with hexane ($3\times10$ mL) and then blown down with nitrogen streams. About 5% aliquot dissolved in toluene was analyzed by GC-FID for quantification, and the remaining sample was analyzed using GC-C-IRMS. With each batch of samples, we prepared analytical blanks to check for potential contamination sources that might have come from solvents and reagents. Internal standards, including FA C~21:0~ and 5-α Cholestane, were added before and after the lipid extraction for quantification.

The GC-MS analysis was performed using an Agilent 7683 B series gas chromatograph attached to an Agilent 5975 inert mass selective detector (Agilent technologies, USA). Samples were injected in splitless mode at 320 °C, and compounds were separated with a VF-17ms column (60 m length, 0.25 mm i.d., 0.25 μm film thickness; J&W Scientific, USA). The carrier gas was helium set at a constant flow rate of 1.3 ml/min. The temperature of the oven was set at 90 °C and held for 1 minute and then was increased at a rate of 5 °C/min to 320 °C, where it was held for 23 minutes. The GC-FID analysis for the FAME samples was carried out using Agilent 6890N gas chromatograph equipped with a PTV inlet in splitless mode (Agilent technologies, USA). The carrier gas was helium flowing at a rate of 1.6 ml/min. The GC system was an Agilent 6890N equipped with a programmed temperature vaporization (PTV) inlet in splitless mode. The GC oven was set at 90°C and held for 1.3 minute, and then was raised to 320°C at a rate of 5°C/min. The final temperature was set for 22 minutes. The GC-C-IRMS analyses of FAMEs of individual fatty acids were conducted using a Thermo Delta V Plus mass spectrometer linked to a GC Trace 1310 through an Isolink Interface (Thermo Fisher Scientific, USA). The sample injection volume was 1 ul in toluene and was injected in splitless mode. Separation was carried out using the same model of column for GC-MS and the GC oven temperature was set at 90 °C for 1.5 minutes and then was raised at a rate of 5 °C/min to 320 °C, with an 18-minute hold. Helium was used as a gas carrier with a constant flow rate of 1.24 ml/minute.

The raw data and R code used for visualizations contained in this paper and R Markdown documents are available online at https://github.com/LiYingWang/GJBpottery to enable research reproducibility and transparency [@Marwick2017; @MarwickandWang2022].

# Results

```{r tidy-data, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/data/raw_data/LY_21_08_20_pot_chrom.CSV"))}
library(tidyverse)
library(ggtext)

# load the GC data
meth <- read_csv(here::here("analysis/data/raw_data/LY_21_08_20_pot_chrom.CSV"))

# data wrangling
ixd <- which(str_detect(meth$Path, "C"))
meths <- split(meth, cumsum(1:nrow(meth) %in% ixd))
meth_data <- meths[c(FALSE, TRUE)]
meth_name <- meths[c(TRUE, FALSE)]
names(meth_data) <- map(meth_data, ~.x$Path[1])
names(meth_name) <-map(meth_name, ~.x$Sample[1])
meths <- map_df(meth_data, ~.x[-1, ] %>%
                   mutate_all(as.numeric), .id = 'sample')

# add sample name
meth_all <-
  meths %>%
  mutate(sample = paste0(parse_number(sample),
                         str_extract(sample, pattern = "_00[0-9]"))) %>%
  mutate(name = case_when(
    sample %in% c("210729_001") ~ "CDG-062",
    sample %in% c("210729_002") ~ "SYG-3",
    sample %in% c("210729_003") ~ "SYG-TN13-E22-2#1",
    sample %in% c("210729_004") ~ "SYG-TN13-E22-1",
    sample %in% c("210729_005") ~ "SYG-TN13-E22-2#2",
    sample %in% c("210729_006") ~ "SYG-TN13-E23-3")) %>%
  mutate(name = factor(name, levels = c("SYG-TN13-E22-1",
                                        "SYG-TN13-E22-2#1",
                                        "SYG-TN13-E22-2#2",
                                        "SYG-TN13-E23-3",
                                        "SYG-3",
                                        "CDG-062")))
```

```{r highlight-compounds, cache=TRUE}
# highlight target compounds
highlight_common <-
  meth_all %>%
  mutate(retention = round(Path, 1)) %>%
  group_by(sample, retention) %>%
  arrange(desc(File)) %>%
  distinct(retention, .keep_all = TRUE) %>%
  mutate(FA = case_when(
    retention == 21.6 ~ "C[14:0]",
    retention == 23.6 ~ "C[15:0]",
    retention == 25.5 ~ "C[16:0]",
    retention == 27.3 ~ "C[17:0]",
    retention == 28.4 ~ "C[18:1]",
    retention == 29.1 ~ "C[18:0]",
    retention == 32.3 ~ "C[20:0]",
    retention == 35.4 ~ "C[22:0]")) %>%
  filter(name %in% c("SYG-3", "SYG-TN13-E22-2#1", "SYG-TN13-E22-1", "SYG-TN13-E23-3", "CDG-062")) %>%
  filter(File > 100000)

highlight_IS <-
  meth_all %>%
  mutate(retention = round(Path, 1)) %>%
  group_by(sample, retention) %>%
  arrange(desc(File)) %>%
  distinct(retention, .keep_all = TRUE) %>%
  mutate(FA = case_when(
    retention == 33.9 ~ "IS",
    retention == 41.3 ~ "IS")) %>%
  filter(name %in% c("SYG-3", "SYG-TN13-E22-2#1", "SYG-TN13-E22-1", "SYG-TN13-E23-3", "CDG-062"))

# highlight the compounds in the potsherd with high concentration
highlight_SYG2_2 <-
  meth_all %>%
  mutate(retention = round(Path, 1)) %>%
  group_by(sample, retention) %>%
  arrange(desc(File)) %>%
  distinct(retention, .keep_all = TRUE) %>%
  mutate(FA = case_when(
    retention == 21.6 ~ "C[14:0]",
    retention == 25.5 ~ "C[16:0]",
    retention == 27.3 ~ "C[17:0]",
    retention == 28.4 ~ "C[18:1]",
    retention == 29.1 ~ "C[18:0]",
    retention == 31.0 ~ "C[19:0]",
    retention == 32.4 ~ "C[20:0]",
    retention == 35.4 ~ "C[22:0]",
    retention == 36.8 ~ "C[23:0]",
    retention == 38.2 ~ "C[24:0]",
    retention == 40.9 ~ "C[26:0]",
    retention == 33.9 ~ "IS",
    retention == 41.3 ~ "IS",
    retention == 45.6 ~ "C[30:0]",
    retention == 47.8 ~ "C[32:0]")) %>%
  filter(name == "SYG-TN13-E22-2#2")
```

```{r chromatogram-four, cache= TRUE}
# plot the first four
chroma_four <-
  meth_all %>%
  filter(name %in% c("SYG-3", "SYG-TN13-E22-2#1", "SYG-TN13-E22-1", 
                     "SYG-TN13-E23-3", "CDG-062")) %>%
  mutate(File = ifelse(File > 1000000, 1000000, File)) %>%
  ggplot(aes(Path, File)) +
  geom_line(size = 0.3) +
  ggrepel::geom_text_repel(data = highlight_common, #ggrepel::geom_text_repel
            aes(label = FA, angle = 0),
            size = 3,
            nudge_y = 120000,
            show.legend = FALSE,
            parse = TRUE) +
  geom_text(data = highlight_IS,
            aes(label = FA),
            size = 3,
            nudge_y = 100000,
            show.legend = FALSE) +
  geom_segment(aes(x= 41.1, xend= 41.5, y= 1000000, yend= 1000000), size =0.3) +
  geom_segment(aes(x= 41.1, xend= 41.5, y= 1030000, yend= 1030000), size =0.3) +
  annotate("text", x = 41.3, y = 600000, label = "IS", size =3) +
  facet_wrap(~name, ncol = 1) +
  scale_y_continuous(labels = scales::comma_format(),
                     limits = c(0, 1100000),
                     breaks = seq(0, 1100000, 200000)) +
  scale_x_continuous(limits = c(20, 45),
                     breaks = seq(20, 45, 5),
                     expand = c(0, 0.5)) + # don't log, many peaks and distort real counts
  labs(x = "retention time", y = "relative intensity") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggsave(plot = chroma_four,
       here::here("analysis","figures", "chromatograms_four.png"),
       bg = "white",
       width = 6,
       height = 8,
       dpi = 300,
       units = "in")
```

The GC-MS analysis for organic residues extracted from potsherds found at the Guijiabao and Gaoshan sites revealed that the amount of lipid compounds ranges from 7 to 48.6 μg/g, above interpretable concentrations (5 μg/g) [@Evershed2008experimental], but were in relatively small amounts in the majority of the potsherds (Table \@ref(tab:FA-composition-table)). The lipid profiles show that the main compounds in most potsherds were saturated fatty acids with an even number of carbons, particularly C~16:0~ (palmitic acid), C~18:0~ (stearic acid) (Figure \@ref(fig:chromatograms-plots-four)). The fatty acids C~16:0~, C~18:0, and C~18:1~ commonly exist in animal fat and plant oil. Unsaturated fatty acid C~18:1~ was also found in all samples, but we are unable to distinguish the C~18:1~ types based on current method. In addition, a number of potsherds had higher contents of C~18:1~ compared to C~18:0~ and C~16:0~, such as SYG-TN13-E23-3.

```{r FA-composition-table, cache= TRUE, cache.extra=tools::md5sum(here::here("analysis/data/raw_data/210824_FAMES_pot_LW_publication.xlsx"))}
quant_FA <-
  readxl::read_excel(here::here("analysis", "data", "raw_data", "210824_FAMES_pot_LW_pub.xlsx"), 
                     skip = 2, col_names = TRUE)

names(quant_FA)[1] <- "number"
names(quant_FA)[2] <- "sample"

quant_FA_number <-
  quant_FA %>%
  pivot_longer(!number&!sample&!`g extracted`&!`%reco`,
               names_to = "FA", 
               values_to = "weight") %>% 
  filter(str_starts(FA, "ng") & !sample =="") %>% 
  mutate(FA = str_remove(FA, "ng "),
         weight = round(weight/`%reco`/`g extracted`/1000, 2)) %>% 
  select(-number, -`%reco`, -`g extracted`) %>% 
  pivot_wider(names_from = sample, 
              values_from = weight)

FA_max_sum <-
  sum(quant_FA_number$"SYG-TN13-E22-2#2")

quant_FA_tidy <-
  quant_FA %>%
  rename("ng total" = "total FAs ug/g") %>%
  pivot_longer(!number&!sample&!`g extracted`&!`%reco`,
               names_to = "FA", 
               values_to = "weight") %>% 
  filter(str_starts(FA, "ng") & !sample =="") %>% 
  filter(!FA == "ng C34:0" & !FA == "ng C21:0") %>% 
  mutate(FA = str_remove(FA, "ng")) %>% 
  mutate(weight = ifelse(str_detect(FA, "C"), 
                         round(weight/`%reco`/`g extracted`/1000, 2), weight)) %>% 
  select(-number, -`%reco`, -`g extracted`) %>% 
  mutate(weight = round(weight, 2),
         weight = ifelse(is.na(weight), "-", weight),
         weight = ifelse(weight == 0, "-", weight)) %>% 
  pivot_wider(names_from = sample, 
              values_from = weight) %>% 
  select("FA", "SYG-TN13-E22-1", "SYG-TN13-E22-2#1", "SYG-TN13-E22-2#2",
         "SYG-TN13-E23-3", "SYG-3", "CDG-062") %>% 
  mutate(FA= str_replace(FA, "C(.*)", "C~\\1~"))

knitr::kable(quant_FA_tidy,
             caption = "Quantitative data of fatty acid (FA) composition (μg/g) for each pottery sample based on GC-FID analysis")
```

```{r chromatograms-plots-four, echo=FALSE, fig.cap= "GC-MS chromatograms of fatty acids extracted from four potsherds from the Guijiabao and Gaoshan sites. 'Cn:x' refers to fatty acid with n carbon atoms and x double bonds. 'IS' indicates internal standard used for quantification, including FA C~21:0~ and 5-α Cholestane"}
knitr::include_graphics(here::here("analysis/figures/chromatograms_four.png"))
```

```{r chromatogram-specific, cache= TRUE}
# plot the specific one
chroma_specific <-
  meth_all %>%
  filter(name %in% c("SYG-TN13-E22-2#2")) %>% 
  ggplot(aes(Path, File)) +
  geom_line(size = 0.3) +
  ggrepel::geom_text_repel(data = highlight_SYG2_2, #ggrepel::geom_text_repel
            aes(label = FA),
            size = 3.5,
            nudge_y = 250000,
            show.legend = FALSE,
            parse = TRUE) +
  scale_y_continuous(labels = scales::comma_format(),
                     limits = c(0, 7500000),
                     breaks = seq(0, 7500000, 1000000)) +
  scale_x_continuous(limits = c(20, 50),
                     breaks = seq(20, 50, 5),
                     expand = c(0, 0.5)) + # don't log, many peaks and distort real counts
  labs(title = "SYG-TN13-E22-2#2", x = "retention time", y = "relative Intensity") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggsave(plot = chroma_specific,
       here::here("analysis","figures", "chromatograms_specific.png"),
       bg = "white",
       width = 8,
       height = 5,
       dpi = 300,
       units = "in")
```

One pottery sample (SYG-TN13-E22-2#2) has much higher lipid concentrations (about `r FA_max_sum` μg/g) than the rest of the samples. A higher yield of lipids might relate to the texture and type of ceramic used or better archaeological contexts for preservation. This sample also has a charred interior surface, indicative of a function associated with food processing. Long-chain fatty acids from C~20:0~ to C~32:0~ are present in that sample, with an even-over-odd carbon chain length predominance, but in lower amounts than in medium-chain components (Figure \@ref(fig:chromatograms-specific)). Although the existence of even-numbered long-chain fatty acids, such as C~30:0~ and C~32:0~, with a high amount of C~16:0~ relative to C~18:0~ might be related to plant waxes [@Post1996; @Copley2005], it is more likely to be associated with degraded animal fat based on a strong predominance of medium-chain fatty acids [@Whelton2018; @Dudd1998]. Overall, the GC-MS results demonstrate strong evidence of degraded animal fats.

```{r chromatograms-specific, echo=FALSE, fig.cap= "The MS chromatogram of the potsherd from the Guijiabao site shows the presence of long-chain fatty acids. 'Cn:x' refers to fatty acid with n carbon atoms and x double bonds. 'IS' indicates internal standards used for quantification, including FA C~21:0~ and 5-α Cholestane"}
knitr::include_graphics(here::here("analysis/figures/chromatograms_specific.png"))
```

```{r isotopic-analysis, cache= TRUE}
library(tidyverse)
isotope_data <-
  readxl::read_excel(here::here("analysis","data", "raw_data","12_21_potshard_13C_pub.xlsx"),
                     sheet = "Summary", col_names =  FALSE)

# filter the Meth corrected data
delta_meth_cor <-
  isotope_data %>%
  select(1, 7, 9) %>%
  slice(2:8) %>%
  dplyr::mutate(...1 = ifelse(is.na(...1), paste0("C13"), ...1))

names(delta_meth_cor) <- delta_meth_cor[1,]

delta_meth_cor <-
  delta_meth_cor %>%
  slice(2:7) %>%
  dplyr::mutate(across(starts_with("13C"), as.numeric)) %>%
  dplyr::mutate(delta = `13C C18:0`-`13C C16:0`) %>%
  mutate(sample = case_when(
    C13 %in% c("Pot 1") ~ "CDG-062",
    C13 %in% c("Pot 2") ~ "SYG-3",
    C13 %in% c("Pot 3") ~ "SYG-TN13-E22-2#1",
    C13 %in% c("Pot 4") ~ "SYG-TN13-E22-1",
    C13 %in% c("Pot 5") ~ "SYG-TN13-E22-2#2",
    C13 %in% c("Pot 6") ~ "SYG-TN13-E23-3")) %>%
  mutate(sample = factor(sample, levels = c("SYG-TN13-E22-1",
                                            "SYG-TN13-E22-2#1",
                                            "SYG-TN13-E22-2#2",
                                            "SYG-TN13-E23-3",
                                            "SYG-3",
                                            "CDG-062"))) %>% 
  mutate(period = ifelse(sample %in% c("SYG-TN13-E22-1", "SYG-3" ),
                         "Historical", "Neolithic"))

min_carbon_isotope <- round(min(delta_meth_cor$`13C C18:0`), 2)
max_carbon_isotope <- round(max(delta_meth_cor$`13C C16:0`), 2)

delta_min <- round(min(delta_meth_cor$delta), 2)
delta_max <- round(max(delta_meth_cor$delta), 2)

E23_C16 <- round(delta_meth_cor$`13C C16:0`[6], 2)
E23_C18 <- round(delta_meth_cor$`13C C18:0`[6], 2)

SYG3_C16 <- round(delta_meth_cor$`13C C16:0`[2], 2)
SYG3_C18 <- round(delta_meth_cor$`13C C18:0`[2], 2)

E22_1_C16 <- round(delta_meth_cor$`13C C16:0`[4], 2)
E22_1_C18 <- round(delta_meth_cor$`13C C18:0`[4], 2)

E22_22_C16 <- round(delta_meth_cor$`13C C16:0`[5], 2)
E22_22_C18 <- round(delta_meth_cor$`13C C18:0`[5], 2)

# plot the carbon isotopes 
isotope_C16_C18 <-
  ggplot(delta_meth_cor, #delta_meth_cor/delta_blank_cor
       aes(`13C C16:0`,`13C C18:0`)) +
  geom_point(size = 2, alpha = 0.9, aes(shape = sample, color = period)) +
  #ggrepel::geom_text_repel(aes(label = C13)) +
  theme_minimal() +
  labs(x = expression(paste(delta*{}^13* C[16:0], "\u2030")),
       y = expression(paste(delta*{}^13* C[18:0], "\u2030"))) +
  scale_x_continuous(limits = c(-33, -20),
                     breaks = seq(-33, -20, 2)) +
  scale_y_continuous(limits = c(-33, -20),
                     breaks = seq(-33, -20, 2)) +
  #coord_fixed(ratio = 1) +
  geom_abline(intercept = -0.7, linetype = "dashed") + #Suryanarayan et al.2021
  geom_abline(intercept = -3.1, linetype = "dashed") +
  annotate("text", x = -21.5, y = -22, angle = 45, vjust = 1.5,
           label = bquote(Delta*{}^13*"C = -0.7 \u2030")) +
  annotate("text", x = -21.5, y = -24.5, angle = 45, vjust = 1.5,
           label = bquote(Delta*{}^13*"C = -3.1 \u2030"))

big_delta_C16 <-
  ggplot(delta_meth_cor, # delta_meth_cor or delta_blank_cor
       aes(`13C C16:0`, delta)) +
  geom_point(size = 2, alpha = 0.9, aes(shape = sample, color = period)) +
  theme_minimal() +
  labs(x = expression(paste(delta*{}^13* C[16:0], "\u2030")),
       y = expression(paste(delta*{}^13* C[18:0], "\u2030"))) +
  scale_x_continuous(limits = c(-33, -20),
                     breaks = seq(-33, -20, 2)) +
  ylim(-2.5, 1.5) +
  geom_hline(yintercept = -0.7, linetype = "dashed") +
  annotate("text", x = -31.5, y = -0.9,
           label = bquote(Delta*{}^13*"C = -0.7 \u2030")) +
  annotate("text", x = -32, y = 1.3,
           label = "C[3]-diet", parse=TRUE) +
  annotate("segment", x = -29, y = 1.3, xend = -24, yend = 1.3,
           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
  annotate("text", x = -21, y = 1.3,
           label = "C[4]-diet", parse=TRUE) +
  annotate("text", x = -22, y = -1.1,
           label = "Ruminant") +
  annotate("text", x = -21.5, y = -0.3,
           label = "Non-ruminant")

library(cowplot)
combined_isotopic_plots <-
  plot_grid(isotope_C16_C18,
          big_delta_C16,
          labels = c('A', 'B'),
          ncol =1,
          align = 'v',
          label_size = 12)

ggsave(here::here("analysis","figures","isotopic_plots.png"),
       bg = "white",
       width = 6,
       height = 8,
       dpi = 300,
       units = "in")
```

To identify the sources of degraded animal fats, we conducted isotope analysis to obtain stable carbon isotope compositions, δ^13^C values of C~16:0~ and C~18:0~, using GC-C-IRMS. The δ^13^C values of C~16:0~ and C~18:0~ have proven useful to distinguish between ruminant adipose, non-ruminant adipose, and ruminant dairy fats based on the different ways dietary carbon sources are utilized by distinct fat types [@Dunne2012; @Copley2003; @Evershed2008]. Considering the bias might be introduced from differential δ^13^C values across geographical regions, we also used the Δ^13^C (δ^13^C~18:0~ - δ^13^C~16:0~) values to determine the sources of degraded fats [@Evershed2008; @Salque2013]. The carbon isotope values for C~16:0~ and C~18:0~ from our samples have a distribution between `r min_carbon_isotope` ‰ and `r max_carbon_isotope` ‰, and their Δ^13^C values range from `r delta_min` ‰ to `r delta_max` ‰. The δ^13^C values were plotted based on the global reference database for the identification of their origins [@Suryanarayan2021; @Dunne2012; @Evershed2008]. Although the referenced isotopic values should be ideally adjusted in terms of local environments, the known reference ranges especially for East Asia could still provide useful information about food categories with cautious comparison.

The results show that the majority of the Neolithic potsherd samples from Guijiabao and Gaoshan contain saturated fatty acids derived from ruminant adipose fats, with one Guijiabao potsherd sample, SYG-TN13-E23-3, indicating a non-ruminant adipose fat origin according to the Δ^13^C value 0.06 ‰ (Figure \@ref(fig:isotopic-plots)). This potsherd (SYG-TN13-E23-3) has the mean δ^13^C values `r E23_C16`‰ and `r E23_C18`‰ for the C~16:0~ and C~18:0~ fatty acids respectively, consistent with the isotopic value ranges of C~3~ plants, terrestrial or freshwater fish (ca. −26.5‰ to -33‰ for both compounds) reported for Neolithic potsherds in southern China [@Shoda2018], and freshwater aquatic products (-27.3‰ to 31.6‰ for the C~16:0~ and -26.4‰ to -29.9 for the C~18:0~) or wild boar (ca. -27‰ to -29.5‰ for both compounds) reported for the Jōmon pottery in western Japan [@Lucquin2016]. However, we are unable to distinguish between the C~3~ plants, terrestrial and freshwater fish by this isotopic approach alone. A further detection of biomarkers for fish such as isoprenoid fatty acids, or for starchy plants and millet will be needed for an better understanding [@Shoda2018; @Lucquin2016; @Heron2016].

For the pottery fragments (SYG-TN13-E22-1 and SYG-3) dating to the historical period, we found that the mean δ^13^C values of the extracts were plotted on the reference line between ruminant adipose and non-ruminant fats. Specifically, SYG-TN13-E22-1 and SYG-3 have the mean δ^13^C values `r E22_1_C16` and `r SYG3_C16` for the C~16:0~ fatty acids, and the mean δ^13^C values `r E22_1_C18` and `r SYG3_C18` for the C~18:0~ fatty acids. It is unsure to know what the origin of these values are, but it is possible that there is a mixing of both non-ruminant sources and ruminant fat [@Copley2003; @Dudd1998milk; @Suryanarayan2021]. It should be noted that the δ13C values could be influenced by multiple different products, such as ruminant products and plant products in the same cooking event, or over different cooking events with different products.

There is only one Neolithic potsherd (SYG-TN13-E22-2#2) with a relatively enriched δ^13^C value (`r E22_22_C16` for the C~16:0~ and `r E22_22_C18` for the C~18:0~) compared with other samples. It is widely known that the δ^13^C values of fatty acids of marine source or C~4~ plants are more enriched than those of ruminant fats [@Lucquin2016]. Given that Guijiabao is situated inland with finding of millet grains, the enriched δ^13^C values might indicate the processing of C~4~ plants instead of marine resources, or the mixture of C~4~ plants and ruminant adipose products due to a slightly depleted δ^13^C value for the C~18:0~ fatty acid. However, in general, most samples have depleted δ^13^C values, which consistent with the δ^13^C values of C~3~ plants or ruminant fats, suggesting the processing of C~3~ plants or C~3~ plants were cooked with ruminant adipose meat. We also could not rule out the possibility that these depleted δ13C values came from mainly ruminant animals fed on C~3~ plants.

```{r isotopic-plots, echo=FALSE, fig.cap= "A: The distribution of stable carbon isotopes for C~16:0~ and C~18:0~. B: The Δ^13^C (δ^13^C~18:0~-~16:0~) values are plotted against δ^13^C~16:0~ values. The dotted lines represent cut-offs for ruminant, non-ruminant, and ruminant dairy fats according to global references"}
knitr::include_graphics(here::here("analysis/figures/isotopic_plots.png"))
```

# Discussion
<!--A brief summary of interesting results-->
This paper is a preliminary study to investigate the pottery use at Guijiabao and Gaushan in Yanyuan area to explore whether those pots were used for cooking and what food categories could be processed in the pots. This will provide a direct evidence of the culinary practices in the Yanyuan region. Although the lipid yields in most of our samples are generally low, interpretative amounts of yields from all of our samples support the assumption that these ceramics were used as cooking or food storage vessels. The low lipid yield could be attributed to the bottom of the vessels, where there were fewer residues than other parts of the vessel, for example, the rim [@Evershed2008experimental; @Kwak2020; @Charters1993]. However, we also found that one sample from the vessel bottom dated to the Neolithic period contains a richer composition of long-chain fatty acids compared with those in other samples. We recognize that there are limitations and challenges for interpretation due to our small sample sizes and a lack of local modern references for animal fats and plant oils. In addition, a detailed analysis of faunal remains from the Guijiabao site is still in progress, which may constrain a comprehensive comparison between different types of evidence. Nevertheless, the lipid profiles and their isotopic ratios could still provide informative clues for a better understanding of culinary practices.

Saturated fatty acids are the major lipid molecules identified in our samples that normally come from animal fats or plant oils, indicative of cooking events. In addition to saturated fatty acids, unsaturated fatty acids such as C~18:1~ was also identified in all samples, and three (SYG-TN13-E22-1, SYG-TN13-E22-2#2, and SYG-TN13-E23-4) out of the six potsherds show a more abundant C~18:1~ than C~18:0~ and C~16:0. Although the specific type of these unsaturated fatty acids can not be further identified in this study without the preparation of specific derivatives [@Berstan2004], their presences might still hint a possibility of plant oils. It is suggested that a higher ratio of C~18:1~ and C~16:0~ to C~18:0~ is associated with a greater likelihood of the origin of plant oils [@Mayyas2018; @Copley2005]. In our samples, the potsherd SYG-TN13_E23-3 has a higher ratio of C~18:1~ and C~16:0~ to C~18:0~, which also corresponds to the category of non-ruminant fats based on the isotopic values. However, we also acknowledge that C18:1 might have originate from animal fats, or there is a potential for modern contamination since monounsaturated fatty acids normally present limited preservation in archaeological contexts [@Whelton2021].

<!--discussion about possible plant origins-->
According to the δ^13^C values of fatty acids, the food residues in Neolithic potsherds could derive from non-ruminant fats, ruminant fats, and mixtures of ruminant fats and C~3~ plants or C~4~ plants. Previous studies suggested that millets (C~4~ plants), including foxtail and broomcorn millets, were the dominant cereals, according to macro-botanical and phytolith studies at Guijiabao [@Guijiabao2016; @Hao2022; @Huan2022]. The processing of millet might also be suggested by one Guijiabao potsherd dating to Neolithic period based on its isotopic ratios. However, a further confirmation would require to identify of any lipid residues associated with C~4~ plants, such as the biomarker for broomcorn millet, miliacin (olean-18-en-3β-ol methyl ether) [@Heron2016]. The non-ruminant fats we identified from a Guijiabao potsherd that might have been from C~3~ plants or freshwater aquatic products, based on similar isotopic values reported by other studies in East Asia [@Shoda2018]. According to the results from flotation analysis at the Guijiabao site, it is possible that the source of C~3~ plants is rice or native wild weeds, such as *Pennisetum* and *Chenopodiaceae*. However, the seeds of wild weeds only account for a small part of the floral assemblage derived from flotation [@Guijiabao2016]. Given that a more depleted isotopic values for 5 out of 6 samples, ~C~3~ plants might consumed more than ~C~4 plants either by local people or by animals that were food source.

We can not rule out other sources for non-ruminant fats, such as freshwater aqautic products, a possibility which can be supported by the finding of fish bones and net sinkers at Guijiabao [@Guijiabao2017b]. This will also need identification of lipid biomarkers related to aquatic oils, such as long-chain $\omega$-(o-alkylphenyl) alkanoic acids and isoprenoid fatty acids [@Evershed2008marine; @Hansel2004; @Cramp2014; @Lucquin2016]. Although horse could be another non-ruminant fat source, whether it was part of local food choices in southwestern China around 4500 years ago during the Neolithic period is doubtful. First, the Pleistocene and Neolithic sites having wild horse bones are primarily concentrated in northern China, with only sporadic sites in the south [@Yuan2006]. Second, domesticated horses first found in southwestern China date to the Dian culture (4th century BC to 1st century AD) in eastern Yunnan [@Lei2009; @Chiou2004], which is 3000 years later than the Neolithic period. Thus, if horse meats is one of the source for non-ruminant fats, then it could only be present in historical ceramics.

<!--discussion about possible ruminant fat origin-->
The ruminant fats identified in the Neolithic potsherds (ca. 4500 BP) from the Guijiabao and Gaoshan sites could have been derived from deer or sheep that were fed on a C~3~ diet, based on their isotopic values that are consistent with global references [@Evershed2002; @Mileto2017]. It is also possible that local wild bovids, such as yaks or water buffalo, were food sources during this period, as these animals have been found in sites in southwestern China [@Zhang2016; @Chen2020]. It is hard to identify the exact ruminant category in this case due to the limited accessibility to local animal bones or fats as reference samples. However, the finding of fatty acid source from ruminant fats, couple with stone artifacts found at the site such as stone balls and arrowheads, suggesting hunting could be an important subsistence activity at the site [@Hao2022; @Guijiabao2020a]. This also corresponds with resources that could be accessed in a highland environment.

We observed mixtures of ruminant fats and C~3~ plants/freshwater aquatic source, based on analyses of two Guijiabao potsherds that date to the historical period (ca. 1000 BP). In addition to rice, the source of C~3~ plants could also have included wheat, since it was introduced into the region during the historical period [@JiangGaoshan2017]. A larger sample size is needed to determine whether there was a change in diet over time. The overall depleted δ^13^C values for potsherd samples associated with ruminant fats suggest that C~3~ plants might have been an important food source for local animals. Despite direct C~4~ plant signals not being found, analysis of one Guijiabao potsherd shows the source of C~4~ plant-fed ruminants during the Neolithic period. This indicates that either C~3~ or C~4~ plants could have been the food source for ruminants, but whether those animals were domesticated is unknown. The zooarchaeological analysis of faunal bones from Guijiabao is still in progress, the results of which could provide more details about local animal species for comparison.

Mixed isotopic signals were also identified from two

Our preliminary results shows that culinary practices in the Neolithic period at Guijiabao include the processing of ruminant adipose products that might be from local wild animals, such as deer. A subsistence that relies on hunting to a certain degree could be possible, given the abundance of wild animals in the natural settings of the Hengduan mountain ranges, as indicated by the faunal remains from the Yingpanshan site [@He2020].

It is also likely that C~3~ plants (eg. rice) was processed based on the isotopic ratios showing non-ruminant fats from one Neolithic potsherds at Guijiabao. In addition to C~3 plants, another possibility for the source of non-ruminant fats could be freshwater aquatic products, which would require confirmation of the presence or absence of aquatic products. It is interesting to detect that one potsherd has a slightly higher values of carbon isotope ratios, indicating a contribution from C4 plants. The detection of C~3~ plants or C~4~ plants might suggest that farming practices were adopted during the time when those crops were first introduced to the Yanyuan area in the Neolithic period. The finding of C~4~ plant signal also corresponds with a previous study that suggests that a C~4~-based diet at the Guijiabao site using isotopic analysis of human bones [@Lin2022]. However, the scale of farming practices and crop consumption is unclear. Despite stone knives and tools that were also found at the sites that might have been related to farming practices, more studies are needed to confirm their usage for harvesting crops [@Zhou2019; @Hao2022; @Lin2022].

An approach analyzing organic residues may only provide a narrow perspective on local diets, but it could improve our current understanding of food practices through presence of food residues that were cooked in ceramic vessels. A large-scale organic residue analysis of ceramics across different pottery forms in a wider region would offer insight into temporal and regional variations in the local diet during the Neolithic period in southwestern China.

An further investigation on the aquatic source should be conducted through selected ion monitoring (SIM) methods in the future to distinguish whether the δ13C
values are the result of processing C3 plants or freshwater aquatic products in the pottery vessel.

sampling bias, poor preservation

<!--- removed
such as C~4~ plants (e.g., millets) and C~3~ plants (e.g., rice), or animals (e.g., pigs) became an essential part of the local diet in Yanyuan during the Neolithic period.
while the food residues in historical potsherds present mixtures of ruminant and C~3~ plants.
-->

# Conclusion
Our case study on organic residues in pottery fragments from the sites at Guijiabao and Gaoshan in southwestern China reveals the presence of lipids that can be analyzed to determine food sources. The foods cooked in the ceramics included ruminant fats and terrestrial plants during the Neolithic period, when millet and rice spread into southwestern China. The source of terrestrial plants could have been C~3~ plants, implying the cooking of rice or other wild weeds in the pots. However, the evidence of direct cooking of either C~3~ or C~4~ plants shows that the amount of C~3~/C~4~ plants cooked is not in proportion to the large quantity of plant remains found at the Guijiabao site, especially C~4~ plants such as millet. Instead, ruminant fats from local animals, such as wild deer, could have been a common food source during the Neolithic period. This suggests that hunting might have been an important part of local subsistence strategy, supplemented by rice/millet farming when they were introduced. Our results from the Guijiabao site could represent a subsistence strategy for adapting to the natural environment of the Hengduan Mountains. We recognize the limitations of our interpretation of the local diet due to the small sample size. Nevertheless, our study could offer an approach using organic residue analysis to determine food choices and explore the extent of the adoption of farming when rice/millet agriculture first spread into the Yanyuan area during the Neolithic period. A more systematic sampling of potsherds across different phases and pottery types may give a clear picture of diet and subsistence in southwestern China during the Neolithic period of transition.

# Acknowledgements
We thank the Chengdu Municipal Institute of Cultural Relics and Archaeology for their permission to access and analyze the pottery samples. We thank Julian Sachs in the Department of Oceanography at the University of Washington for his support in letting us use their laboratory for sample preparation and for conducting lipid extraction and analysis. We thank Matthew Wolhowe for his suggestions and assistance with the GC-based machines. We also want to express our gratitude to the anonymous reviewers for their comments that helped us improve this paper. This research is sponsored by the National Science and Technology Council (MOST 109-2410-H-001-009-MY2).

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```