### 2024-01-21 analysis/paper/paper.Rmd
substantial edits

---
title: "Preliminary insights into the use of pottery and culinary pratices at Guijiabao site in southwestern China"
author:
  - Li-Ying Wang:
      email: liyingw@asihp.net
      institute: [AcS]
      correspondence: true
  - Kuei-chen Lin:
      email: kueichenlin@asihp.net
      institute: [AcS]
      correspondence: false
  - Zhiqing Zhou:
      email: zhouzhiQ2000@163.com
      institute: [InsC]
      correspondence: false
institute:
  - AcS: Institute of History and Philology, Academia Sinica 
  - InsC: Chengdu Institute of Cultural Relics and Archaeology
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |

     The emergence of mixed agriculture of millet and rice in Neolithic southwestern China has been a topic of extensive discussion. Despite the recovery of both millet and rice grains, the varying proportions observed across the area underscore the complexity of the process of agricutural dispersal. Further investigation is needed to understand how new crops were integrated into southwestern China, characterized by diverse environmental conditions and landscapes. This study contributes to this understanding by exploring subsistence strategies in southwestern Sichuan through a preliminary analysis of lipid residues in pottery from the Guijiabao site using organic geochemical analyses. Situated in the middle Hengduan mountain ranges, the Guijiabao site presents well-preserved stratigraphic records spanning from the Neolithic to the historical period, providing valuable insights into local subsistence practices in a highland environment. Results reveal the detection of ruminant adipose fats in the majority of potsherds, with a few showing plant inputs such as broomcorn millet. This suggests that animals played a substantial role in the local diet, alongside the consumption of millet. These culinary practices also indicate that hunting may have remained a significant subsistence activity in the high-altitude region of southwestern China, even after the introduction of crops.

keywords: |
  Southwestern China; Neolithic period; Lipid residue analysis; Compound-specific carbon isotope analysis; Pottery; Culinary practices
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300)

image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
```

# Introduction

<!-- intro to the topic and research aim-->
The movement of Neolithic populations has long been discussed in archaeology to understand the processes of agricultural diffusion, regional variations, and their consequences [@Bellwood2023; @Davison2006]. In southwest China, the emergence of farming practices was thought to be associated with the migrating population from central or northern China in the Neolithic period [@Deng2019]. Crop remains found in prehistoric southwest China, including foxtail millet, broomcorn millet, and rice, can be traced back to as early as 5000 BP in southwest Sichuan and around 4600 BP in western Yunnan [@Jiang2016; @Deng2019; @YunnanMuseum1981; @Zhang2010; @Dal2022]. Despite different views on the routes of agricultural dispersal [@Guedes2013; @Guedes2018; @Liu2022; @Huan2022], it is generally believed that mixed farming was the main agricultural practice in southwest China [@Deng2019]. However, more studies are still needed to understand how migrant communities interacted with new environments, especially in an area with a high diversity of landscapes, such as high mountains, plateaus, basins, and terraces.

<!--Guijaibao as a research area-->
This study focuses mainly on archaeological sites located in the Hengduan Mountain ranges, a region in southwest China connecting the Tibetan Plateau to the northwest, the Sichuan Basin to the east, and the Yunnan Plateau to the south. As a critical passage for population movement, the Hengduan Mountain ranges and valleys facilitate not only material exchange [@Tong1986; @Hein2014; @Hein2016; @Ma2021], but also the spread of agriculture even further to mainland Southeast Asia since the Neolithic period [@Liu2022; @Stevens2017]. The Guijiabao site in Yanyuan Basin, the middle Hengduan Mountain ranges, are a key site with the earliest Neolithic human occupation in this region dating to 5000 B.P., continuing to the historical period [@Guijiabao2017a; @Guijiabao2017b; @Huan2022; @Hao2022]. The complete stratigraphy at Guijiabao provides a unique opportunity to investigate questions relative to subsistence since the Neolithic period in a basin surrounded by high mountains. For example, what is the role of agriculture in local environments? What other resources were obtained in addition to crops? In this study, we present our preliminary analyses of archaeological potsherds from the Guijiabao site and the Gaoshan site in southwest China using organic residue analysis of lipids.

<!--justifying organic residue analysis-->
Although the remains of animal bones and plants offer evidence of subsistence, their preservation heavily depends on environmental conditions. Geographical regions with seasonal heavy precipitation and acidic sediments, such as Southwest Sichuan and the middle Hunduan Mountain [@Li1989], tend to lead to poor preservation of archaeological remains. In such cases, organic residue analysis is a powerful tool for studying ancient food practices by directly examining residues in pottery. Organic residues, specifically lipids, can be well preserved in the porous fabric of pottery for thousands of years due to the protection of clay matrices from microbiological degradation and water leaching [@Evershed1993; @Evershed2008; @Roffet2017; @Dudd1999; @Miller2020]. The growing application of organic residue analysis demonstrates its usefulness in understanding ancient pottery use [@Kimpe2004; @Mayyas2022], human subsistence strategies [@Kwak2020], and culinary practices across a region over time [@Craig2011; @Suryanarayan2021; @Zhang2022; @Junno2020].

<!--regional application-->
By identifying lipid compositions and conducting compound-specific carbon isotope analysis, archaeologists are able to distinguish between plant oils, ruminant fats, dairy products, and aquatic/marine resources [@Copley2005; @Craig2007; @Evershed1994]. Recent studies have successfully identified various plant types processed in archaeological vessels, including broomcorn millet [@Bossard2013; @Heron2016; @Ganzarolli2018], starchy plants [@Shoda2018], and cereals [@Hammann2018; @Colonese2017]. The increasing use of ceramics in China highlights the robustness of organic residue analysis for comprehending regional cooking practices [@Shoda2018; @Lyu2024; @Han2022], discerning distinct subsistence strategies between low- and high-altitude zones [@Zhang2022], and exploring the impact of nomadic pastoralism on northwest China [@Sun2023]. In this study, we aim to investigate pottery use and culinary practices in the middle Hengduan mountains, focusing on ceramics from the Guijiabao and Gaoshan sites to understand local subsistence.

# Geographical and Cultural Background of Guijiabao in Yanyuan Basin

```{r SW-china-map, fig.cap = "The archaeological sites Guijiabao and Gaoshan and other locations in southwestern China that are mentioned in this study. Map data is from naturalearthdata.com"}
library(here)
knitr::include_graphics(here("analysis", "figures", "Sichuan-sites-map.png"))
```

<!--Yanyuan area and Guijiabao site-->
The Guijiabao site is located in the southwest of Yanyuan Basin, a pull-apart basin in the middle eastern part of the Hengduan Mountain ranges (Figure \@ref(fig:SW-china-map)), with an elevation of 2400 meters. The first survey at Guijiabao in 2015 and subsequent excavations from 2016 to 2019 revealed abundant archaeological materials and features, such as semi-subterranean and ground houses, post holes, ash pits, ditches, and burials, suggesting a settlement site [@Hao2022; @Guijiabao2017a]. The site has been occupied since the Neolithic period (5000-3700 BP), continued to the Bronze Age (3200-2700 BP) and the historical period (1200-500 BP) corresponding to the late North Song to early Ming dynasties, based on radiocarbon dates [@Hao2022; @Huan2022]. The Neolithic period can be further divided into two subphases, 5000-4500 BP and 4500-3700 BP, according to the artifact assemblages and ceramic styles.

<!--the major finding at GJB-->
Ceramics are the major finding at Guijiabao. The common vessel forms include jars, urns, bowls, and basins. The representative ceramic styles in the Neolithic period are grayish high-tempered fine ware and reddish brown coarse ware with cord-marked or basket-impressed patterns [@Hao2022; @Guijiabao2017a]. Ceramics with charred surfaces suggest that they might be used for cooking. This research also attempts to confirm that these pottery vessels were used as cooking or storage wares. Some ceramic decorations present northern styles, while others with incised geometric patterns and impressed dot patterns resemble those found in western Yunnan, indicative of close relationships or interaction between Guijiabao and neighboring areas, the northern and southern parts of the Hengduan Mountains [@Guijiabao2017a]. The flotation analysis of soil samples from the 2015 excavation demonstrates the predominance of cultivated broomcorn millet (*Panicum miliaceum*) and foxtail millet (*Setaria italica*) at Guijiabao during the Neolithic period [@Guijiabao2016]. Rice grains (*Oryza sativa*) were also identified in later excavations from the Neolithic period based on macro botanical evidence and phytolith analysis [@Hao2022; @Huan2022], though in a smaller number than millet.

<!--info about Gaoshan-->
To compare the culinary practices in different ecological zones, we examined a sample from the Gaoshan site in southwestern Chengdu Plain, an alluvial plain in Sichuan close to the Tibetan Plateau (Figure \@ref(fig:SW-china-map)). Excavations at the Gaoshan site reveal abundant materials, including ceramics, stone artifacts, animal and plant remains, and archaeological features such as house foundations, walls, burials, ash pits, ditches, and a well [@Gaoshan2017; @He2020; @JiangGaoshan2017; @ZhouGaoshan2015]. According to similar pottery forms and styles, the Gaoshan site can be dated to the early Baodun period, the earliest Neolithic culture (ca. 4700-3700 BP) in the Chengdu Plain [@Gaoshan2017]. The primary subsistence strategy at the Gaoshan site could have been rice cultivation with pig husbandry, supplemented by millet cultivation [@JiangGaoshan2017; @Lee2020; @He2020]. It is suggested that the spread of rice farming into Sichuan occurred at least 4700 years ago through population migration from the Middle and Lower Yangzi valleys [@Zhang2010].

While mixed agriculture of millet and rice was practiced in southwestern China, variations in their proportions across the region have been observed. Macro-botanical and phytolith analyses of the Guijiabao site indicate that millet was the primary crop, aligning with findings at the Yingpanshan site in Sichuan [@Zhao2011]. Conversely, botanical evidence from numerous sites in the Chengdu plain suggests a prevalence of rice, as seen at the Gaoshan and Baodun sites [@JiangGaoshan2017], and the Baiyangcun site in Yunnan [@Jiang2016; @Henglanshan2014]. Moreover, while most research focuses on agricultural practices, there is a lack of clarity regarding other resources utilized by early farmers in new environments. This study aims to address this gap by examining direct evidence from pottery residues, using geochemical techniques to provide insights into culinary practices through a case study in southwest China.

# Materials and methods

## Sampling

As a preliminary study to assess the feasibility of employing organic geochemical approaches, we collected five pottery fragments from cultural layers in the residential area of Guijiabao (sample code starting with SYG) (Table \@ref(tab:sherds-table)). For comparison, one fragment was sampled from Goashan (sample code starting with CDG). Those pottery fragments represent typical types with well-defined contexts, with their radiocarbon dates ranging from the Neolithic period (ca. 4500 BP) to the historical period (ca. 1000 BP). The pottery samples are mostly base fragments and fine sand-tempered, with a few displaying impressed decorations. By examining potsherds from the early and late periods at Guijiabao, we aim to explore pottery use across time periods and the potential of utilizing an organic geochemical approach.

```{r sherds-table, cache= TRUE, cache.extra=tools::md5sum(here::here("analysis/data/raw_data/SW-china-potsherd.xlsx"))}
Ch_sherds <-
  readxl::read_excel(here::here("analysis", "data", "raw_data", "SW-china-potsherd.xlsx"), col_names = TRUE)

knitr::kable(Ch_sherds,
             caption = "A list of the pottery sherds that were analyzed in this study. The sherd labeled 'SYG' indicates potsherds from the Guijiabao site, while the label 'CDG' refers to those from the Gaoshan site")
```

## Lipid extraction

The surface of each pottery sample was removed to prevent external contamination from the environment and handling. Approximately 3.5 grams of powdered samples were extracted with a mixture of dichloromethane/methanol solution (9:1 v/v) using accelerated solvent extraction (ASE Dionex 350; at 100 °C) [@Mcclure2018]. To obtain a higher yield, free fatty acids were recovered from bound lipids through saponification, which was conducted by adding 10 mL of 1M KOH in methanol and 1 mL water to the extracts and heating at 70 °C overnight [@Kaluzna2016; @Craig2004]. Saponified extracts were acidified with 6M hydrochloric acid (HCl), and the acid fraction was extracted using hexane ($3\times10$ mL). A 5% aliquot of the extract was treated with 20 uL N,O-bis(trimethylsilyl)trifluoroacetamide (BSTFA) containing 1% v/v trimethylchlorosilane (TMCS), and heated at 70 °C for one hour. After drying under a stream of nitrogen, the derivatized sample was dissolved in toluene for gas chromatography mass spectrometry (GC-MS).

The remaining 95% aliquot was fractioned using solid phase extraction (Aminopropyl, NH2, column with 0.5 g sorbent) with 8 mL dichloromethane:Isopropyl alcohol (3:1), 8 mL 4% acetic acid in ethyl ether, and 6 mL methanol. The second fraction was methylated to produce fatty acid methyl esters (FAMEs) with 1 mL dry hexane and 2 mL 10% (v/v) acetyl chloride in dry methanol, and heated at 70 °C overnight. The FAME derivatives were extracted with hexane ($3\times10$ mL). About a 5% aliquot was analyzed by gas chromatography with flame-ionization detection (GC-FID) for quantification, and the remaining sample was analyzed using gas chromatography combustion isotope ratio mass spectrometry (GC-C-IRMS) [@Craig2007; @Evershed1994]. Analytical blanks were prepared and analyzed using the same procedure as the pottery samples to monitor potential contamination sources from solvents and reagents. Internal standards, FA C~21:0~ and 5-α Cholestane or n-tetratriacontane, were added before and after the lipid extraction.

## GC methods
The GC-MS analysis was performed using an Agilent 7683 B series gas chromatograph attached to an Agilent 5975 inert mass selective detector (Agilent technologies, USA). Samples were injected in splitless mode at 320 °C. Compounds were separated with a VF-17ms column (60 m length, 0.25 mm i.d., 0.25 μm film thickness; J&W Scientific, USA). The carrier gas was helium flowing at a rate of 1.3 ml/min. The oven temperature was set at 90 °C for 1 minute, then increased at a rate of 5 °C/min to 320 °C and held for 23 minutes. In addition to scan mode, a selected ion monitoring (SIM) mode was used to further detect miliacin (olean-18-en-3β-ol methyl ether; *m/z* 189, 204, 231, 425, 440) [@Heron2016]. The oven temperature was set at 50 °C for 1 min, then raised at a rate of 10°C/min to 310 °C and held for 15 min. An authentic standard of miliacin was also run to monitor the retention time.

The GC-FID analysis of FAMEs was carried out using an Agilent 6890N gas chromatograph equipped with a PTV inlet in splitless mode (Agilent technologies, USA). The carrier gas was helium flowing at a rate of 1.6 ml/min. The GC system was an Agilent 6890N equipped with a programmed temperature vaporization (PTV) inlet in splitless mode. The GC oven was set at 90°C for 1.3 minute, then raised at a rate of 5°C/min to 320°C and held for 22 minutes. The GC-C-IRMS analyses of FAMEs of palmitic (C~16:0~) and stearic (C~18:0~) were conducted using a Thermo Delta V Plus mass spectrometer linked to a GC Trace 1310 through an Isolink Interface (Thermo Fisher Scientific, USA) in splitless mode. Separation was carried out using the same model of column for GC-MS. Helium was used as a gas carrier with a constant flow rate of 1.24 ml/minute. The oven temperature was set at 90 °C for 1.5 minutes, then raised at a rate of 5 °C/min to 320 °C and held for 18 minutes.

The data generated by these analyses and R code used for visualizations are openly available at https://github.com/LiYingWang/GJBpottery to enable research reproducibility and transparency [@Marwick2017; @MarwickandWang2022].

# Results

```{r tidy-data, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/data/raw_data/LY_21_08_20_pot_chrom.CSV"))}
library(tidyverse)
library(ggtext)

# load the GC data
meth <- read_csv(here::here("analysis/data/raw_data/LY_21_08_20_pot_chrom.CSV"))

# data wrangling
ixd <- which(str_detect(meth$Path, "C"))
meths <- split(meth, cumsum(1:nrow(meth) %in% ixd))
meth_data <- meths[c(FALSE, TRUE)]
meth_name <- meths[c(TRUE, FALSE)]
names(meth_data) <- map(meth_data, ~.x$Path[1])
names(meth_name) <-map(meth_name, ~.x$Sample[1])
meths <- map_df(meth_data, ~.x[-1, ] %>%
                   mutate_all(as.numeric), .id = 'sample')

# add sample name
meth_all <-
  meths %>%
  mutate(sample = paste0(parse_number(sample),
                         str_extract(sample, pattern = "_00[0-9]"))) %>%
  mutate(name = case_when(sample %in% c("210729_001") ~ "CDG-062",
                          sample %in% c("210729_002") ~ "SYG-3",
                          sample %in% c("210729_003") ~ "SYG-TN13-E22-2#1",
                          sample %in% c("210729_004") ~ "SYG-TN13-E22-1",
                          sample %in% c("210729_005") ~ "SYG-TN13-E22-2#2",
                          sample %in% c("210729_006") ~ "SYG-TN13-E23-3")) %>%
  mutate(name = factor(name, levels = c("SYG-TN13-E22-1",
                                        "SYG-TN13-E22-2#1",
                                        "SYG-TN13-E22-2#2",
                                        "SYG-TN13-E23-3",
                                        "SYG-3",
                                        "CDG-062")))
```

```{r FA-composition-tidy, cache= TRUE, cache.extra=tools::md5sum(here::here("analysis/data/raw_data/210824_FAMES_pot_LW_publication.xlsx"))}
quant_FA <-
  readxl::read_excel(here::here("analysis", "data", "raw_data", "210824_FAMES_pot_LW_pub.xlsx"), 
                     skip = 2, col_names = TRUE)

names(quant_FA)[1] <- "number"
names(quant_FA)[2] <- "sample"

quant_FA_number <-
  quant_FA %>%
  pivot_longer(!number&!sample&!`g extracted`&!`%reco`,
               names_to = "FA", 
               values_to = "weight") %>% 
  filter(str_starts(FA, "ng") & !sample =="") %>% 
  mutate(FA = str_remove(FA, "ng "),
         weight = round(weight/`%reco`/`g extracted`/1000, 2)) %>% 
  select(-number, -`%reco`, -`g extracted`) %>% 
  pivot_wider(names_from = sample, 
              values_from = weight) %>% 
  filter(!FA %in%c("C34:0", "C21:0"))

FA_sum_min <- sum(quant_FA_number$"SYG-TN13-E22-2#1", na.rm = T)
FA_sum_max <- sum(quant_FA_number$"SYG-TN13-E22-2#2")
```

The total lipid concentrations of potsherds at Guijiabao and Gaoshan range from `r FA_sum_min` to `r FA_sum_max` μg/g (Table \@ref(tab:FA-composition-table)). Lipid yields from all potsherds are above interpretable concentrations (5 μg/g) [@Evershed2008experimental], suggesting those ceramics should be associated with food processing. The main compounds were saturated fatty acids with an even number of carbons, particularly C~16:0~ (palmitic acid) and C~18:0~ (stearic acid), commonly derived from animal fats and plant oils (Figure \@ref(fig:chromatograms-plots-four)). Unsaturated fatty acid, C~18:1~, was detected in all samples and dominant in three, but we are unable to distinguish the C~18:1~ types based on current methods. The most interesting finding is that milliacin, the compound specific to broomcorn millet [@Heron2016; @Bossard2013; @Ganzarolli2018], was identified in one Neolithic sample (SYG-TN13-E22-2#2) from Guijiabao (Figure \@ref(fig:chromatograms-specific-std)). This sample has a wider range of fatty acids from C~12:0~ to C~30:0~ (`r FA_sum_max` μg/g), where long-chain fatty acids could originate from plant sources [@Post1996; @Copley2005]. Overall, the dominance of C~16:0~ and C~18:0~ in most samples suggests that the lipids extracted from our samples were greatly degraded.

```{r FA-composition-table, cache= TRUE}
quant_FA_tidy <-
  quant_FA %>%
  rename("ng total" = "total FAs ug/g") %>%
  pivot_longer(!number&!sample&!`g extracted`&!`%reco`,
               names_to = "FA", 
               values_to = "weight") %>% 
  filter(str_starts(FA, "ng") & !sample =="") %>% 
  filter(!FA == "ng C34:0" & !FA == "ng C21:0") %>% 
  mutate(FA = str_remove(FA, "ng")) %>% 
  mutate(weight = ifelse(str_detect(FA, "C"), 
                         round(weight/`%reco`/`g extracted`/1000, 2), weight)) %>% 
  select(-number, -`%reco`, -`g extracted`) %>% 
  mutate(weight = round(weight, 2),
         weight = ifelse(is.na(weight), "-", weight),
         weight = ifelse(weight == 0, "-", weight)) %>% 
  pivot_wider(names_from = sample, 
              values_from = weight) %>% 
  select("FA", "SYG-TN13-E22-1", "SYG-TN13-E22-2#1", "SYG-TN13-E22-2#2",
         "SYG-TN13-E23-3", "SYG-3", "CDG-062") %>% 
  mutate(FA= str_replace(FA, "C(.*)", "C~\\1~"))

knitr::kable(quant_FA_tidy,
             caption = "Quantitative data of fatty acid (FA) composition (μg/g) for each pottery sample based on GC-FID analysis")
```

```{r highlight-compounds, cache=TRUE}
# round all peaks in order
all_peaks <-
  meth_all %>%
  mutate(retention = round(Path, 1)) %>%
  group_by(sample, retention) %>%
  arrange(desc(File)) %>%
  distinct(retention, .keep_all = TRUE)

# highlight target compounds
highlight_common <-
  all_peaks %>% 
  mutate(FA = case_when(retention == 21.6 ~ "C[14:0]",
                        retention == 23.6 ~ "C[15:0]",
                        retention == 25.5 ~ "C[16:0]",
                        retention == 27.3 ~ "C[17:0]",
                        retention == 28.4 ~ "C[18:1]",
                        retention == 29.1 ~ "C[18:0]",
                        retention == 32.3 ~ "C[20:0]",
                        retention == 35.4 ~ "C[22:0]")) %>%
  filter(name %in% c("SYG-3", "SYG-TN13-E22-2#1", "SYG-TN13-E22-1", "SYG-TN13-E23-3", "CDG-062")) %>%
  filter(File > 100000)

highlight_IS <-
  all_peaks %>% 
  mutate(FA = case_when(retention == 33.9 ~ "IS",
                        retention == 41.3 ~ "IS")) %>%
  filter(name %in% c("SYG-3", "SYG-TN13-E22-2#1", "SYG-TN13-E22-1", "SYG-TN13-E23-3", "CDG-062"))
```

```{r chromatogram-four, cache= TRUE}
# to insert ticks in plot
insert_minor <- function(major_labs, n_minor) {labs <-
  c( sapply( major_labs, function(x) c(x, rep("", 4) ) ) )
labs[1:(length(labs)-n_minor)]}

# plot the first group
chroma_four <-
  meth_all %>%
  filter(name %in% c("SYG-3","SYG-TN13-E22-2#1","SYG-TN13-E22-1","SYG-TN13-E23-3","CDG-062")) %>%
  mutate(File = ifelse(File > 1000000, 1000000, File)) %>%
  ggplot(aes(Path, File)) +
  geom_line(size = 0.3) +
  ggrepel::geom_text_repel(data = highlight_common, #ggrepel::geom_text_repel
            aes(label = FA, angle = 0),
            size = 3,
            nudge_y = 115000,
            show.legend = FALSE,
            parse = TRUE) +
  geom_text(data = highlight_IS,
            aes(label = FA),
            size = 3,
            nudge_y = 100000,
            show.legend = FALSE) +
  geom_segment(aes(x= 41.1, xend= 41.5, y= 1000000, yend= 1000000), size =0.3) +
  geom_segment(aes(x= 41.1, xend= 41.5, y= 1030000, yend= 1030000), size =0.3) +
  annotate("text", x = 41.3, y = 1100000, label = "IS", size =3) +
  facet_wrap(~name, ncol = 1) +
  scale_y_continuous(limits = c(0, 1160000),
                     breaks = seq(0, 1100000, 200000),
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(18, 47),
                     breaks = seq(15, 50, 1),
                     labels = insert_minor(seq(15, 50, 5), 4),
                     expand = c(0, 0)) +
  labs(x = "Retention time (mins)", y = expression("Relative abundance " %->% "")) +
  theme_classic() +
  theme(axis.line.y = element_blank(), strip.background = element_blank(), # remove title box
        axis.ticks.y=element_blank(), axis.text.y = element_blank())

ggsave(plot = chroma_four,
       here::here("analysis","figures", "chromatograms_four.png"),
       bg = "white",
       width = 6,
       height = 8,
       dpi = 300,
       units = "in")
```

```{r chromatograms-plots-four, echo=FALSE, fig.cap= "Partial total ion cuurent (TIC) chromatograms of pottery samples from Guijiabao and Gaoshan. 'Cn:x' refers to fatty acid with n carbon atoms and x double bonds. 'IS' indicates internal standard used for quantification, including FA C~21:0~ and 5-α Cholestane"}

knitr::include_graphics(here::here("analysis/figures/chromatograms_four.png"))
```

```{r chromatogram-specific, cache= TRUE}
## miliacin
miliacin <- read_csv(here::here("analysis/data/raw_data/LY_miliacin_std_chrom.CSV"))
ixd_3 <- which(str_detect(miliacin$Path, "D"))
meths_3 <- split(miliacin, cumsum(1:nrow(miliacin) %in% ixd))
meth_data_3 <- meths_3[c(FALSE, TRUE)]
meth_name_3 <- meths_3[c(TRUE, FALSE)]
names(meth_data_3) <- map(meth_data_3, ~.x$Path[1])
names(meth_name_3) <-map(meth_name_3, ~.x$Sample[1])
meths_3 <- map_df(meth_data_3, ~.x[-1, ] %>%
                    mutate_all(as.numeric), .id = 'sample')

highlight_m <-
  meths_3  %>%
  mutate(retention = round(Path, 1)) %>%
  group_by(sample, retention) %>%
  arrange(desc(File)) %>%
  distinct(retention, .keep_all = TRUE) %>%
  mutate(std = case_when(retention == 29.4 ~ "M", retention == 30.5 ~ "IS"))

std_miliacin <-
  meths_3%>%
  ggplot(aes(Path, File)) +
  geom_line(size = 0.3) +
  ggrepel::geom_text_repel(data = highlight_m, #ggrepel::geom_text_repel
                           aes(label = std, angle = 0), size = 3.5, nudge_y = 100000,
                           show.legend = FALSE, parse = TRUE) +
  scale_y_continuous(limits = c(0, 4600000),
                     breaks = seq(0, 4600000, 1000000),
                     expand = c(0,0)) +
  scale_x_continuous(limits = c(13.5, 32.5),
                     labels = insert_minor(seq(10, 35, 5), 4),
                     breaks = seq(10, 35, 1),
                     expand = c(0, 0)) +
  labs(x = "Retention time (mins)", y = expression("Relative abundance " %->% "")) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), axis.line.y = element_blank(),
        axis.ticks.y=element_blank(), axis.text.y = element_blank())

# load data produced by program 2
meth_SYG22 <- read_csv(here::here("analysis/data/raw_data/LY_pot_SYG_E22_22_chrom.CSV"))

# tidy data
ixd_2 <- which(str_detect(meth_SYG22$Path, "D"))
meths_2 <- split(meth_SYG22, cumsum(1:nrow(meth_SYG22) %in% ixd))
meth_data_2 <- meths_2[c(FALSE, TRUE)]
meth_name_2 <- meths_2[c(TRUE, FALSE)]
names(meth_data_2) <- map(meth_data_2, ~.x$Path[1])
names(meth_name_2) <-map(meth_name_2, ~.x$Sample[1])
meths_2 <- map_df(meth_data_2, ~.x[-1, ] %>%
                  mutate_all(as.numeric), .id = 'sample')

# peaks in SYG-E22-2#2
peaks_SYGE22 <-
  meths_2 %>%
  mutate(retention = round(Path, 1)) %>%
  group_by(sample, retention) %>%
  arrange(desc(File)) %>%
  distinct(retention, .keep_all = TRUE)

highlight_SYG_FA <-
  peaks_SYGE22 %>%
  mutate(FA = case_when(retention == 17.7 ~ "C[15:0]",
                        retention == 18.7 ~ "C[16:0]",
                        retention == 19.7 ~ "C[17:0]",
                        retention == 20.4 ~ "C[18:1]",
                        retention == 20.6 ~ "C[18:0]",
                        retention == 21.5 ~ "C[19:0]",
                        retention == 22.4 ~ "C[20:0]",
                        retention == 23.3 ~ "IS",
                        retention == 24.0 ~ "C[22:0]",
                        retention == 24.8 ~ "C[23:0]",
                        retention == 26.3 ~ "C[25:0]",
                        retention == 27.0 ~ "C[26:0]",
                        retention == 28.3 ~ "C[28:0]",
                        retention == 30.5~ "IS"))

highlight_std_small <-
  peaks_SYGE22 %>%
  mutate(FA = case_when(retention == 14.2 ~ "C[12:0]",
                        retention == 15.4 ~ "C[13:0]",
                        retention == 16.6 ~ "C[14:0]",
                        retention == 25.6 ~ "C[24:0]",
                        retention == 29.4 ~ "M"))

## plot the one with high yield (GC program 2)
chroma_specific_2 <-
  meths_2 %>%
  ggplot(aes(Path, File)) +
  geom_line(size = 0.3) +
  ggrepel::geom_text_repel(data = highlight_SYG_FA , #ggrepel::geom_text_repel
                           aes(label = FA, angle = 0), size = 3.5, nudge_y = 100000,
                           show.legend = FALSE, parse = TRUE) +
  ggrepel::geom_text_repel(data = highlight_std_small, #ggrepel::geom_text_repel
                           aes(label = FA),
                           size = 3.5,
                           segment.alpha = .25,
                           nudge_y = 320000, #min.segment.length = 300000,
                           show.legend = FALSE,
                           parse = TRUE) + #
  scale_y_continuous(labels = scales::comma_format(),
                     limits = c(0, 4600000),
                     breaks = seq(0, 4600000, 1000000),
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(13.5, 32.5),
                     labels = insert_minor(seq(10, 35, 5), 4),
                     breaks = seq(10, 35, 1),
                     expand = c(0, 0)) +
  labs(x = "Retention time (mins)", y = expression("Relative abundance " %->% "")) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), axis.line.y = element_blank(),
        axis.ticks.y=element_blank(), axis.text.y = element_blank())

library(cowplot)
SYG_with_miliacin <-
  plot_grid(chroma_specific_2, std_miliacin ,
            labels = c('A', 'B'), ncol =1, align = 'v', label_size = 12)

ggsave(plot = SYG_with_miliacin,
       here::here("analysis","figures", "chromatograms_specific_std.png"),
       bg = "white", width = 6, height = 8, dpi = 300, units = "in")
```

```{r chromatograms-specific-std, echo=FALSE, fig.cap= "A: Partial total ion current (TIC) chromatogram of a Guijiabao sample, SYG-TN13-E22-2#2, showing the presence of long-chain fatty acids and miliacin. Cn:x refers to fatty acid with n carbon atoms and x double bonds; IS indicates internal standards (FA C~21:0~ and n-tetratriacontane); M refers to miliacin. B: Partial TIC chromatogram of miliacin (olean-18-en-3β-ol ME) with n-tetratriacontane as internal standard"}
knitr::include_graphics(here::here("analysis/figures/chromatograms_specific_std.png"))
```

```{r isotopic-analysis, cache= TRUE}
library(tidyverse)
isotope_data <-
  readxl::read_excel(here::here("analysis","data", "raw_data","12_21_potshard_13C_pub.xlsx"),
                     sheet = "Summary", col_names =  FALSE)

# filter the Meth corrected data
delta_meth_cor <-
  isotope_data %>%
  select(1, 7, 9) %>%
  slice(2:8) %>%
  dplyr::mutate(...1 = ifelse(is.na(...1), paste0("C13"), ...1)) %>% 
  janitor::row_to_names(1) %>% 
  dplyr::mutate(across(starts_with("13C"), as.numeric))

delta_meth_data <-
  delta_meth_cor %>%
  dplyr::mutate(delta = `13C C18:0`-`13C C16:0`) %>%
  mutate(name = case_when(C13 %in% c("Pot 1") ~ "CDG-062",
                          C13 %in% c("Pot 2") ~ "SYG-3",
                          C13 %in% c("Pot 3") ~ "SYG-TN13-E22-2#1",
                          C13 %in% c("Pot 4") ~ "SYG-TN13-E22-1",
                          C13 %in% c("Pot 5") ~ "SYG-TN13-E22-2#2",
                          C13 %in% c("Pot 6") ~ "SYG-TN13-E23-3")) %>%
  mutate(name = factor(name, levels = c("SYG-TN13-E22-1",
                                        "SYG-TN13-E22-2#1",
                                        "SYG-TN13-E22-2#2",
                                        "SYG-TN13-E23-3",
                                        "SYG-3",
                                        "CDG-062"))) %>%
  mutate(period = ifelse(name %in% c("SYG-TN13-E22-1", "SYG-3" ), "Historical", "Neolithic"))

rg_C16_isotope <- round(c(min(delta_meth_data$`13C C16:0`), max(delta_meth_data$`13C C16:0`)), 2)
rg_C18_isotope <- round(c(min(delta_meth_data$`13C C18:0`), max(delta_meth_data$`13C C18:0`)), 2)

delta_range <- round(c(min(delta_meth_data$delta), max(delta_meth_data$delta)), 2)

SYG3_isotope <- round(c(delta_meth_data$`13C C16:0`[2], delta_meth_data$`13C C18:0`[2]), 2)
E22_1_isotope <- round(c(delta_meth_data$`13C C16:0`[4], delta_meth_data$`13C C18:0`[4]), 2)
E22_22_isotope <- round(c(delta_meth_data$`13C C16:0`[5], delta_meth_data$`13C C18:0`[5],
delta_meth_data$delta[5]), 2)
E22_21_isotope <- round(c(delta_meth_data$`13C C16:0`[3], delta_meth_data$`13C C18:0`[3], delta_meth_data$delta[3]), 2)
E23_isotope <- round(c(delta_meth_data$`13C C16:0`[6], delta_meth_data$`13C C18:0`[6], delta_meth_data$delta[6]), 2)

library(png)
tem3 <- readPNG(here::here("analysis","figures", "isotope_range.png"), native = FALSE, info = FALSE)

# plot the carbon isotopes 
isotope_C16_C18 <-
  delta_meth_data %>% #delta_meth_cor/delta_blank_cor
  ggplot(aes(`13C C16:0`,`13C C18:0`)) +
  geom_point(size = 2, alpha = 0.9, aes(shape = name, color = period)) +
  #ggrepel::geom_text_repel(aes(label = C13)) +
  labs(x = expression(paste(delta*{}^13* C[16:0], "\u2030")),
       y = expression(paste(delta*{}^13* C[18:0], "\u2030"))) +
  scale_x_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +
  scale_y_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +#coord_fixed(ratio = 1)
  geom_abline(intercept = -0.7, linetype = "dashed") + #Suryanarayan et al.2021
  geom_abline(intercept = -3.1, linetype = "dashed") +
  annotate("text", x = -21.5, y = -22, angle = 45, vjust = 1.5,
           label = bquote(Delta*{}^13*"C = -0.7 \u2030")) +
  annotate("text", x = -21.5, y = -24.5, angle = 45, vjust = 1.5,
           label = bquote(Delta*{}^13*"C = -3.1 \u2030")) +
  annotation_raster(tem3, ymin = -38, ymax= -17.9, xmin = -36, xmax = -18.3) +
  theme_minimal()

big_delta_C16 <-
  delta_meth_data %>%
  ggplot(aes(`13C C16:0`, delta)) +
  geom_point(size = 2, alpha = 0.9, aes(shape = name, color = period)) +
  labs(x = expression(paste(delta*{}^13* C[16:0], "\u2030")),
       y = expression(paste(delta*{}^13* C[18:0], "\u2030"))) +
  scale_x_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +
  ylim(-3, 1.5) +
  geom_hline(yintercept = -0.7, linetype = "dashed") +
  annotate("text", x = -31.5, y = -0.9,
           label = bquote(Delta*{}^13*"C = -0.7 \u2030")) +
  annotate("text", x = c(-32, -21), y = c(1.3, 1.3),
           label = c("C[3]~diet", "C[4]~diet"), parse= TRUE) +
  annotate("segment", x = -29, y = 1.3, xend = -24, yend = 1.3,
           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
  annotate("text", x = c(-22, -21.5), y = c(-1.1, -0.3),
           label = c("Ruminant", "Non-ruminant")) +
  theme_minimal()

library(cowplot)
combined_isotopic_plots <-
  plot_grid(isotope_C16_C18, big_delta_C16,
          labels = c('A', 'B'), ncol =1, align = 'v', label_size = 12)

ggsave(here::here("analysis","figures","isotopic_plots.png"),
       bg = "white", width = 6, height = 8, dpi = 300, units = "in")
```

To identify the sources of degraded lipids, we conducted compound-specific carbon isotope analysis of C~16:0~ and C~18:0~ using GC-C-IRMS. This technique has proven useful to distinguish fat origin between ruminant, non-ruminant, ruminant dairy products, C~3~ plants, C~4~ plants, and marine resources based on their stable carbon isotope values (hereafter δ13C ) [@Dunne2012; @Evershed2008]. For example, the δ13C values from marine sources or C~4~ plants are more enriched than those from ruminant adipose fat and C~3~ plants due to different metabolic or photosynthetic pathways [@Copley2003; @Meier2002; @Cramp2014]. The values of δ^13^C~16:0~ and δ^13^C~18:0~ were plotted based on global modern references for the assessment of their origin. To minimize potential bias introduced from differential δ^13^C values across geographical regions, the values of δ^13^C~16:0~ and Δ^13^C (δ^13^C~18:0~ - δ^13^C~16:0~) were also plotted for identification of animal fats [@Copley2003; @Salque2013]. Although it is hard to identify the exact food category in this case due to the limited accessibility to local animal bones or fats as reference samples, references from neighboring areas with cautious comparison still allow meaningful interpretation.

The isotope data shows that δ^13^C~16:0~ values range from `r rg_C16_isotope[1]` ‰ to `r rg_C16_isotope[2]` ‰, δ^13^C~18:0~ values range from `r rg_C18_isotope[1]` ‰ to `r rg_C18_isotope[2]`, and Δ^13^C values are between `r delta_range[2]` ‰ and `r delta_range[1]` ‰. A half of the samples (n= 3) at Guijiabao and Gaushan fall within the global reference range for ruminant adipose fat, where one sample from Guijiabao (SYG-TN13-E22-2#2) has a more enriched δ^13^C~16:0~ (`r E22_22_isotope[1]`) than others. The enriched δ^13^C suggests that the diet of ruminants contains both C~4~ and C~3~ plants (Figure \@ref(fig:isotopic-plots): B), while the depleted δ^13^C is related to a more C~3~ plant diet. While marine resources could potentially contribute to enriched δ^13^C values, the inland location of Guijiabao, situated approximately 1000 km away from the current South China Sea coast, restricts this possibility.

One sample (SYG-TN13-E23-3) with Δ^13^C of `r E23_isotope[3]` falls within the range for non-ruminant fats (Figure \@ref(fig:isotopic-plots): B). Its relatively depleted δ^13^C~ values (δ^13^C~16:0~ of `r E23_isotope[1]` ‰ and δ^13^C~18:0~ of `r `E23_isotope[2]` ‰) reflect input of C~3~ plants, terrestrial animals, or freshwater resources [@Shoda2018; @Lucquin2016]. The two samples (SYG-TN13-E22-1 and SYG-3) dating to the historical period clustered within a narrow range of δ^13^C~16:0~ (`r SYG3_isotope[1]` ‰ to `r E22_1_isotope[1]`) and δ^13^C~18:0~ (`r SYG3_isotope[2]` ‰ to `r E22_1_isotope[2]`) near the reference line, showing a mixing source of ruminant and non-ruminant fat [@Copley2003]. A mixture of lipids from different sources is possible because δ13C values could be influenced by multiple distinct products, such as ruminant fats and plants, in the same cooking event or separate cooking events over time [@Evershed2002].

```{r isotopic-plots, echo=FALSE, fig.cap= "A: The distribution of stable carbon isotopes for C~16:0~ and C~18:0~. The ellipses represent modern animal references fed on a C~3~diet (Copley et al. 2003). B: The Δ^13^C (δ^13^C~18:0~-~16:0~) values are plotted against δ^13^C~16:0~ values. The dash lines represent cut-offs for ruminant, non-ruminant, and ruminant dairy fats based on Suryanarayan et al. 2021"}

knitr::include_graphics(here::here("analysis/figures/isotopic_plots.png"))
```

# Discussion
<!--A brief summary of the lipid yield and function of pottery-->
Lipids extracted from potsherds at Guijiabao and Gaoshan all exceed interpretative amounts, with their profiles revealing the processing of both animal and plant products This confirms the role of the pottery as vessels for cooking or food storage at both sites. The generally low lipid concentrations (<20 μg/g) in five out of the six samples may be attributed to the bottom parts of vessels, which might absorb fewer residues compared to other parts, such as the rim, during cooking events [@Evershed2008experimental; @Kwak2020; @Charters1993]. Additional factors contributing to the low lipid yields could be post-depositional processes [@Correa2014], or diverse methods of food preparation, including grilling or air-drying [@Zhang2022]. We recognize challenges for interpreting our results due to the small sample size and the absence of a direct local reference database from faunal bones or modern animals. Furthermore, the detailed contexts of faunal remains at Guijiabao are still in progress, thus they are not fully accessible. These limitations constrain a thorough comparison of various types of evidence at the site. Despite these restrictions, the overall pattern of lipid profiles with isotopic values offers valuable insights into culinary practices.

<!--discussion about possible ruminant fat origin-->
Neolithic potsherds from Guijiabao and Gaushan (ca. 4500 BP) reveal food contents, including ruminant adipose fat (n= 3) and non-ruminant fat (n= 1). The ruminant adipose (Δ^13^C from `r E22_21_isotope[3]` to `r E22_22_isotope [3]` ) could have originated from goat/sheep, cattle, and deer based on a modern database [@Evershed2002], alongside archaeological contexts in northwestern China [@Sun2023]. At Guijiabao, the ruminants' diet includes C~3~ plants (SYG-TN13-E22-2#1), and a mixture of C~3~ and C~4~ plants (SYG-TN13-E22-2#2). A C~3~ diet is supported by the predominance of C~3~ weeds at Guijiabao, such as *Pennisetum* and *Chenopodiaceae* [@Guijiabao2016]. Some species in *Chenopodiaceae* can undergo C~4~ photosynthesis in arid environments [@Rudov2020; @Chasan2022], which may explain the mixture of plant types in SYG-TN13-E22-2#2. The direct input from C~4~ plants could be another reason for the mixed signal, as we have identified millet in the same sample. The domestication status of these ruminants detected in the samples remains unclear. However, considering the finding of hunting tools at Guijiabao, such as stone balls and arrowheads [@Hao2022; @Guijiabao2020a], with the abundance of wild ruminants in the natural settings of the Hengduan mountain ranges, such as deer, mutton, and yaks [@Zhang2022; @Zhang2016; @Chen2020; @He2020], it is plausible that subsistence at Guijiabao involving hunting to a certain degree.

<!--Plant as possible non-ruminant fat origin-->
The non-ruminant fat (SYG-TN13-E23-3) could originate from C~3~ plants and freshwater resources, based on similar isotopic ranges reported for Neolithic pottery in eastern China [@Shoda2018; @Lyu2024], and a global database [@Steele2010]. Although fish bones and net sinkers were found at Guijiabao [@Guijiabao2017b], lipid molecular evidence for fish, such as ω-(o-alkylphenyl)alkanoic acids and isoprenoid fatty acids [@Cramp2014], was absence in our samples. This suggests that fish might not have been commonly consumed, or alternative cooking methods for fish were employed. Possible source C~3~ plants could be rice or wild weeds based on botanical evidence [@Guijiabao2016; @Hao2022; @Huan2022]. Due to a small portion of seeds of wild weeds in the floral assemblage [@Guijiabao2016], coupled with rice phytolichs identified from the residential area [@Huan2022], rice appears to be the more likely major source of C~3~ plants. In addition, the non-ruminant sample has a higher ratio of C~18:1~ and C~16:0~ to C~18:0~, associated with a greater likelihood of plant oils [@Mayyas2018; @Copley2005]. However, we acknowledge that C18:1 might also originate from animal fats, or there is a potential for modern contamination, as monounsaturated fatty acids typically have limited preservation in archaeological contexts [@Whelton2021]. A further detection of biomarkers for starchy plants with a specific derivatization method for lipid compounds will be necessary to assess the presence of rice oil [@Shoda2018].

<!--Wild non-ruminant as possible non-ruminant fat origin-->
The possibility of terrestrial animal input for the non-ruminant fat, such as wild boar, cannot be ruled out. The isotopic values of the non-ruminant sample align with those of reference fats from modern wild boars in western Japan [@Lucquin2016] and in the southeast Tibetan Plateau [@Zhang2022]. Although horses could be another non-ruminant fat source, it is less likely to have been part of the local diet in southwestern China during the Neolithic period. This is because wild horse bones are primarily concentrated at the Pleistocene and Neolithic sites in northern China, with only sporadic occurrences in the south [@Yuan2006]. Moreover, substantial evidence of domesticated horses in southwestern China dates back to the Dian culture (4th century BC to 1st century AD) in eastern Yunnan [@Lei2009; @Chiou2004], a period approximately 3000 years later than the Neolithic period. Ongoing research on faunal remains at Guijiabao will provide more details for further discussion.

<!--ruminants and crops-->
Our preliminary data reveals the predominance of ruminants fed on a C~3~ and C~3~/C~4~ diet in pottery vessels during the Neolithic period, suggesting a potentially more significant role of ruminants in local culinary practices than previously thought. Most studies have primarily focused on mixed farming of millet and rice as the main economy in southwest China after the introduction of agricultural in the Neolithic period [@Deng2019]. At Guijaibao, domesticated crops, especially broomcorn millet, constitute the majority of plants based on macro- and micro- botanical analyses [@Guijiabao2016; @Huan2022], and nitrogen and carbon isotope analysis of human bones suggests a more C~4~ diet [@Lin2022]. Additionally, stone knives found at the site might be potentially be used for farming practices [@Zhou2019; @Hao2022; @Lin2022]. In our study, a Neolithic potsherd containing the lipid compound of broomcorn millet [@Heron2016], provides direct evidence of millet processing at Guijiabao. Although a depleted δ^13^C~16:0~ in the non-ruminant sample suggest the possibility of C~3~plants related to rice, further analysis is needed for confirmation. It is important to note that lipid signatures of plants tend to be weak due to microbial degradation or being masked by animal fats [@Colonese2017; @Evershed2008experimental], possibly explaining the low lipid yield of plants in one sample. Nevertheless, the frequent processing of ruminant meats in pottery, with unearthed hunting tools, suggests that animal hunting could be an essential subsistence strategy alongside crop consumption. This may also reflect a subsistence strategy adapted to highlands with cold environments where meat consumption is common [@Zhang2022].

The two historical potsherds (ca. 1000 BP) from Guijiabao demonstrates mixtures of ruminant adipose and non-ruminant fats, implying a potential increase in the exploitation of C3 plants, pigs, or freshwater resources compared to the Neolithic period. In addition to rice, the source of C~3~ plants could also have included wheat, since it was introduced into the region during the historical period [@JiangGaoshan2017]. For the Gaoshan sample, a slightly enriched δ^13^C may suggest minor contributions of pigs [@Evershed2002], in correspondence with the predominance of domestic pigs at the site [@He2020]. However, determining changes in culinary practices over time and regional differences would require a larger sample size. While a geochemical approach analyzing organic residues may provide a limited perspective on local diets, it contributes to improving our current understanding of direct food practices through organic residues in ceramic vessels. As a preliminary study, our data reveals food contents cooked in pottery from both animal and plant sources, offering valuable insights into culinary practices during the time when agriculture spread into the Yuan Basin in southwest China.

# Conclusion
This paper presents a preliminary result investigating pottery use at Guijiabao in the Yanyuan area, middle Hengduan mountain range, through organic residue analyses to explore culinary practices. Despite the small sample size, our results reveal a relatively diverse range of food categories cooked in pottery during the Neolithic period, including ruminant adipose fats, plants such as broomcorn millet, and non-ruminant sources. In the historical period, mixtures of ruminant adipose fats and non-ruminant fats were identified. Non-ruminant fats could have originated from C3 plants, pigs, or freshwater resources.

While a mixed crop of millet and rice is generally accepted as the main subsistence in southwest China, our results indicate that ruminant adipose fats were more commonly cooked in pottery. The prevalence of ruminant fats suggests the significance of meat consumption, a aspect often overlooked after the introduction of crops. The discovery of hunting tools at Guijiabao further indicates that hunting might have played a crucial role in the local subsistence strategy, complementing rice/millet farming. TThis provides valuable insights into subsistence strategies in a highland setting in the Hengduan Mountains ranges. Although our preliminary data has limitations for a comprehensive interpretation, the results demonstrate the feasibility of lipid identifications with compound-specific stable isotope analysis to determine food content processed in ceramic vessels at Guijiabao. A more systematic sampling of potsherds across different phases and pottery types will contribute to a clearer understanding of diet and subsistence in southwestern China in the future.

# Acknowledgements
We thank the Chengdu Municipal Institute of Cultural Relics and Archaeology for their permission to access and analyze the pottery samples. We thank Julian Sachs in the Department of Oceanography at the University of Washington for his support in letting us use their laboratory for sample preparation and for conducting lipid extraction and analysis. We thank Matthew Wolhowe for his suggestions and assistance with the GC-based machines. We also want to express our gratitude to the anonymous reviewers for their comments that helped us improve this paper. The research was supported by the National Science and Technology Council (MOST 109-2410-H-001-009-MY2).

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```